---
title: "Fitbit_measures"
author: "Bachir"
date: "7 February 2019"
output: html_document
---

# *Reproductive Research - Course Project 1*

## Activity monitoring

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### *Loading and preprocessing the data*
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="./Factivity.zip",method="auto")  
unzip(zipfile = "Factivity.zip")
unlink("Factivity.zip")
```

### *Load and read the data*
```{r}
library(readr)
data1 <- read_csv("activity.csv")
head(data1)
```

### *Process the data*
```{r}
library(plyr)
```

# 1. Mean total number of steps taken per day

## Total number of steps taken per day
```{r}
totalstepsperday <- with(data1, aggregate(steps, list(date), sum, na.rm = TRUE))
names(totalstepsperday) <- c("date", "steps")
```

## Histogram of the total number of steps taken each day
```{r}
with(totalstepsperday, hist(steps, col = "pink",
                          xlab = "Steps per day", main = "Total number of steps taken each day"))
```

### Mean and median of the total number of steps taken per day
```{r}
moy1 <- mean(totalstepsperday$steps, na.rm = TRUE)
med1 <- median(totalstepsperday$steps, na.rm = TRUE)
```
The average steps taken per day is `r moy1` and the median is `r med1`.

# 2. Average daily activity pattern
```{r}
totatstepsperint <- with(data1, aggregate(steps, list(interval), mean, na.rm = TRUE))
names(totatstepsperint) <- c("interval", "steps")
```

## Plot of the 5-minute interval and the average number of steps taken
```{r}
with(totatstepsperint, plot(interval, steps, type = "l", lwd = 2,
                            xlab = "5-minute interval", ylab = "Average number of steps",
                            main = "Average number of steps taken across all days"))
```

## The 5-minute interval that contains the maximum number of steps?
```{r}
maxstep <- with(totatstepsperint, interval[which.max(steps)])
```
The maximum number of steps for 5-minute interval is `r maxstep`.

# 3. Imputing missing values

## Total number of missing values in the dataset
```{r}
miss <- with(data1, sum(is.na(steps)))
```
The total number of missing values in the dataset is `r miss`.

## Fill the missing values with the mean for each 5-minutes interval
```{r}
data2 <- data1
NA.position <- which(is.na(data2$steps))

for (i in 1:length(NA.position)) {
  data2[NA.position[i],1] <- with(data2, aggregate(steps,
                                                   list(interval == data2[NA.position[i],3][[1]]),
                                                   mean, na.rm = TRUE))[2,2]
                                 }

sum(is.na(data2$steps)) # check
```

### *Total number of steps taken per day*
```{r}
totalstepsperday2 <- with(data2, aggregate(steps, list(date), sum))
names(totalstepsperday2) <- c("date", "steps")
```

## Histogram of the total number of steps taken each day
```{r}
with(totalstepsperday2, hist(steps, col = "pink",
                            xlab = "Steps per day", main = "Total number of steps taken each day"))
```

### Mean and median of the total number of steps taken per day
```{r}
moy2 <- mean(totalstepsperday2$steps)
med2 <- median(totalstepsperday2$steps)
```
After filling the missing data, the average steps taken per day becomes `r moy2` and the median is `r med2`.

# 4. Differences in activity patterns between weekdays and weekends

### *Creation of a factor variable for "weekday" and "weekend"*
```{r}
data3 <- data2
data3$date <- as.Date(data3$date, format="%Y-%m-%d")
data3 <- mutate(data3, week = ifelse(weekdays(data3$date) == "Saturday" |
                                               weekdays(data3$date) == "Sunday", "weekend", "weekday"))
data3$week <- as.factor(data3$week)
```

### *Average weekly activity pattern*
```{r}
totatstepsperint2 <- with(data3, aggregate(steps, list(interval, week), mean))
names(totatstepsperint2) <- c("interval", "week", "steps")
```

## Panel plot of the 5-minute interval and the average number of steps taken
```{r}
library(ggplot2)
g <- ggplot(totatstepsperint2, aes(interval, steps))
g + geom_line() + facet_grid(week ~ .)
```
